\section{2018-07-31 Lecture}

The mathematics of imaging

What is an inverse problem?

Some examples here

Impedance tomography

Apply a voltage to the boundary of a region, measure current and try to determine conductivity.

Let $\Omega \subseteq \RR^n$ be an open bounded set with a smooth boundary.
We will model conductivity with a (smooth) symmetric matrix $\gamma \in \cC^\infty(\Omega,\Sym)$ which is also bounded: $C > \gamma(x)> 1/C$ for all $x \in \bar\Omega$ for some $C$.
This accounts for different conductivities in different directions.
The input voltage (potential) is a function $f \in \cC^\infty(\p\Omega)$.
Let $u$ be the interior induced voltage (potential).
Then Kirchoff's laws state that
\[ \nabla \cdot \gamma \nabla u_f = 0 \]
in $\Omega$, where $u_f|_{\p\Omega} = f$ (a Dirichlet condition), and $\hat n \cdot \gamma \nabla u_f|_{\p\Omega} \in \cC^\infty(\p\Omega)$ is the surface flux (a Neumann condition).
(Here $\gamma\nabla u_f$ is the current and $\hat n$ is the normal vector at the boundary.)

How do we know a solution exists?
Leave this question for now.

We have a Dirichlet-Neumann map (also known as a voltage-current map):
\[ \Lambda_\gamma: f \mapsto \hat n \cdot \gamma \nabla u_f \in \cC^\infty(\p\Omega). \]
Is $\gamma \mapsto \Lambda_\gamma$ injective?
NO!
If $F: \Omega \to \Omega$ is smooth,  $F\inv$ is smooth and $F|_{\p\Omega}=\id$, then $\Lambda_\gamma = \Lambda_{\tilde\gamma}$ where
\[ \tilde\gamma = \left( \frac{dF \, \gamma \, dF^\intercal}{\abs{\det dF}} \right). \]
This is just a change of basis.

\begin{exer}
  Peove this.
  Hint: $\nabla \cdot \gamma \nabla u = 0 \iff \nabla \cdot \tilde \gamma \nabla \tilde u = 0$ for $\tilde u = u \circ F\inv$.
\end{exer}

It is conjectured that this is the only obstruction (i.e.\@ if $\Lambda_{\gamma_1}=\Lambda_{\gamma_2}$ then there exists an $F: \Omega \to \Omega$ smooth with smooth inverse such that $\gamma_1 = F_* \gamma_2$.
This is known in dimension 2.

We will look at a simpler problem

Assume that $\vec\gamma(x) = \gamma(x) \id$ for $\gamma(x) \in \cC^\infty(\Omega,\RR)$.
(So the conductivity is isotropic, hence scalar.)
Then $\gamma \mapsto \Lambda_\gamma$ is injective.

This is known as the \textbf{Calder\'on problem}.

Course outline:

\begin{itm}
  \io
  We will solve the Calder\'on problem for $n \geq 3$.
  (The case $n=2$ is harder and requires another trick.)

  \io
  The partial data problem for $n \geq 3$ (when the boundary conditions are known only over a subset of the whole boundary).
  Can we still solve the Calder\'on problem?

  \io
  The case $n=2$.

  \io
  Fundamental theory.
  We will use theorems without proof in the earlier sections.
  So this course is wil proceed from the top to the bottom.
\end{itm}
