\section{2018-11-01 Lecture}

Stefanov-Yang Yang 2018

Let $(M,g)$ be a Lorentzian manifold.
For simplicity, let $g$ be the symmetric $2$-tensor
\[ g = -dt^2 + h_{jk} \, dx^j \; dx^k \]
where $h_{jk}$ depends on $x$ and $t$, and we set $t=x^0$.
Then the \textbf{d'Alembertian} operator is
\begin{align*}
  \square_g &= \frac{1}{\sqrt{\abs{\det g}}} \p_j\left( \sqrt{\abs{\det g}} g^{jk} \p_k u \right) \\
  &= -\frac{1}{\sqrt{\det h}} \p_j\left( \sqrt{\det h} \p_t u \right) = \frac{1}{\sqrt{\det h}} \p_j\left( \sqrt{\det h} h^{jk} \p_k u \right) \\
\end{align*}
This is the analog of the wave operator for Lorentzian manifolds.

Define a \textbf{causal structure} on the manifold.
Let $p\in M$ and $v\in T_pM$.
Then $v$ is \textbf{time-like} (or \textbf{null}) if $g(v,v)<0$, \textbf{light-like} if $g(v,v)=0$ and \textbf{space-like} if $g(v,v)>0$.
A hyperspace $S$ is called time-like if for all $p\in S$, $T_pS$ contains a light-like vector.
Suppose $M$ is light-like.

A time-like or null vector is \textbf{future pointing} if $g(v,\p t)<0$.

\begin{lem}
  Let $v'\in T_p(\p M)$ be time-like and future pointing.
  Then there exists a unique future pointing null vector $v$ such that $v-v'\perp T_pM$ with respect to $g$, and $v$ points in to $M$.
\end{lem}

Dynamics of null-geodesics (i.e.\@ light)
Let $(p_0,v_0')\in T_pM$ be time-like and future pointing.
Let $v_0$ be the unique future pointing null vector in the Lemma above.
Let $(\gamma_{p_0,v_0'})$ be the null geodesic originating from $p_0$ in direction $v_0$ (so the velocity vector is always a null vector).
Now flow along this geodesic until it reaches a point $p_1$ on $\p M$.
If the velocity at $p_1$ is $v_1$, then we project to the light-like subspace to give $v_1$.
Then $\cL_g:(p_0,v_0') \mapsto (p_1,v_1')$ is known as the \textbf{lens relation}.

PDE and wavefront.

Let $f\in\cD'(\p M)$ and with $\supp f$ contained in a neighbourhood of $p_0$ in $\p M$ and consider $\WF(f)$ in a small neighbourhood of $(p_0,v_0')$.
Also let $q\in\cC^\infty(M)$.
For all such $f$, there exists a unique $u$ (up to a $\cC^\infty$ function) solving
\begin{equation*}
  \begin{cases}
    \left( \square_g+q \right)u \in \cC^\infty \\
    (u-f)|_{\p M} \in \cC^\infty,
  \end{cases}
\end{equation*}
where $\WF(u)$ contains only future-pointing null vectors.
This is known as a microlocal solution.

Let $\Gamma_0$ be a small strip-shaped neighbourhood about the geodesic $\gamma_{p_0,v_0'}$.
WLOG we can assume $\supp u\subseteq\Gamma_0$.
Define $\Lambda_{g,q}f=(\p_\nu u)|_{\p M}$.

\begin{thm}[Stefanov-Yang Yang]
  If $\Lambda_{g_1,q_1}=\Lambda_{g_2,q_2}$, then $\cL_{g_1}=\cL_{g_2}$, and
  \[ \int_\gamma q_1-q_2=0 \]
  for all null future-pointing geodesics.
\end{thm}

\begin{proof}
  SY showed that
  \[ \left( \Lambda_{g_1,q_1}f \right)(x) = \int_{\p M} \int_{\theta\in\RR^N} e^{i\ph(x,y,\theta)} a(x,y,\theta) f(y) \, d\theta \, dy. \]
  Recall from last lecture
  \[ \Lambda_\ph = \left\{ (x,d_x\ph,y,d_y\ph) \mid d_\theta\ph(x,y,\theta)=0 \right\}. \]
  SY showed that
  \[ \Lambda_\ph = \left\{ (x,v_0',\cL(x,v_0')) \right\}. \qedhere \]
\end{proof}
