\section{2018-03-05 Lecture}

intro

textbook allen hatcher

email lecturer to tell him you are in the class

assignment is already up

exercises in lecture notes on website for checking understanding

exercises in book are harder

final exam problems similar to book

topology

take a topological space and assign to it some topological structure (like a group) which is invariant according to some allowed manipulations
it also gave rise to homological algebra which is used in representation theory, category theory and so on

chapter 0: introduction
1. basic notions of topology
1.1 topological spaces

\begin{defn}
	A topological space is a pair $(X,\tau)$, where $X$ is a set and $\tau$ is a topology for the set:
	\begin{itemize}
		\item $\emptyset, X \in \tau$
		\item $\{V_\alpha \in \tau\} \implies \bigcup_\alpha V_\alpha \in \tau$
		\item $\{V_\alpha \in \tau\} \implies \bigcap_\alpha V_\alpha \in \tau$ for finite sets
	\end{itemize}
\end{defn}

\begin{exam}
	Euclidean space $\RR^n$ with distance function $d: \RR^n \times \RR^n \to \RR$ with
	\[d(x,y) = \sqrt{\sum_{i=1}^n(x_i-y_i)^2}\]
	\[\]
	Whenever we use Euclidean space we assume this topology. fill in
	
	open ball
	open set condition
\end{exam}

\begin{defn}
	Given a topological space $(X,\tau)$, any subset $S \subset X$ is a topological space for $\tau_S = \{S \cap U \mid U \in \tau\}$ (the subspace topology).
\end{defn}

We will sometimes omit the $\tau$ when talking about maps.

\begin{exam}
	\begin{itemize}
		\item $I = [0,1] \subset \RR$ fill in
	\end{itemize}
	Closed Ball (n)
	Sphere (n-1)	
\end{exam}

1.2 continuous maps and homeomorphisms

\begin{defn}
	A map $f: X \to Y$ of sets is continuous if $f\inv = \{x \in X | f(x) \in V\}$ is open for all $V \subset Y$ open.
\end{defn}

Continuous maps are the natural maps between topological spaces.
If we talk about a map between topological spaces it is assumed to be continuous.

\begin{exam}
	$S \subset X$ fill in
\end{exam}

composition of maps is continuous

category top of topological spaces

when are two spaces the same topologically

\begin{defn}
	A map $f: X \to Y$ is a homeomorphism if
	\begin{itemize}
		\item $f$ is continuous
		\item $f$ is bijective
		\item $f\inv$ is continuous
	\end{itemize}
	Alternatively, a homeomorphism is a bijection of sets where $f[U]$ is open in $Y$ if and only if $U$ is open in $X$.
	If a homeomorphism exists then we write $X \cong Y$.
\end{defn}

In topology we consider to be the same some spaces which would not be considered the same in analysis or geometry.

\begin{exam}
	The unit square is isomorphic to the unit disc.
	\[D^n \cong \square^n = \{x \in \RR^n | -1 \leq x_i \mid \leq 1\}\]
	The boundary of $\square^n$ is not smooth but this does not matter in topology.
\end{exam}

1.3 Elementary operations on topological spaces

Let $X$ and $Y$ be topological spaces.

\begin{defn}
	The product space $X \times Y$ is:
	\begin{itemize}
		\item the set $X \times Y = \{(x,y) \mid x \in X, y \in Y\}$, and where
		\item open sets in $X \times Y$ are the unions of $U \times V$ where $U$ is open in $X$ and $V$ is open in $Y$
	\end{itemize}
\end{defn}

We have (continuous) maps
\begin{itemize}
	\item $P_X: X \times Y \to X, (x,y) \mapsto x$
	\item $P_X: X \times Y \to Y, (x,y) \mapsto y$
\end{itemize}

\begin{exer}
	Let $f: Z \to X \times Y$.
	Then $f$ is continuous if and only if $P_Xf$ and $P_Yf$ is continuous.
\end{exer}

\begin{exam}
	\item $\RR \times \RR \cong \RR^2$
	\item $S^1 \times S^1 \cong T^2$. (sphere times sphere is torus)
\end{exam}

\begin{defn}
	Take an equivalence relation (reflexive, symmetric and transitive) $~$ on $X$.
	Define $Z \defeq X/~$ be the set of equivalence classes with quotient map. fill in fix tilde
\end{defn}

\begin{exam}
	quotient by a subspace
	Let $S \subset X$.
	Then for $x \neq y \in X$, let $x~y$ if and only if $x,y \in S$.
	Then $X/S \defeq X/~$.
\end{exam}

\begin{defn}
	Consider the disjoint union of sets $X \sqcup Y$.
	The open subsets are $U \sqcup V$ for $U$ open in $X$ and $V$ open in $Y$.
	We have inclusion maps fill in
\end{defn}
