\section{2018-03-05 Lecture}

chapter 0: introduction
1. basic notions of topology
1.1 topological spaces

\begin{defn}
	A \textbf{topological space} is a pair $(X,\tau)$, where $X$ is a set and $\tau$ is a \textbf{topology} for the set:
	\begin{itm}
		\item $\emptyset, X \in \tau$
		\item $\{V_\alpha \in \tau\} \implies \bigcup_\alpha V_\alpha \in \tau$
		\item $\{V_\alpha \in \tau\} \implies \bigcap_\alpha V_\alpha \in \tau$ for finite sets
	\end{itm}
\end{defn}

\begin{exam}
	An example of a topological space is Euclidean space $\RR^n$ with distance function $d: \RR^n \times \RR^n \to \RR$ with
	\[d(x,y) = \sqrt{\sum_{i=1}^n(x_i-y_i)^2}\]
	\[\]
	Whenever we use Euclidean space we assume this topology.
	Define
	\[\Dd(x,r) = \{y \in \RR^m \mid d(x,y)<r\}\]
	for $x \in \RR^n$ and $r \in \RR$ positive.
	Then a set $U \subset \RR^n$ is defined to be open if for all $x \in U$ there exists an $\epsilon>0$ such that $\Dd(x,\epsilon) \subset U$.
\end{exam}

\begin{defn}
	Given a topological space $(X,\tau)$, any subset $S \subset X$ is a topological space for $\tau_S = \{S \cap U \mid U \in \tau\}$ (the \textbf{subspace topology}).
\end{defn}

We will sometimes omit the $\tau$ when talking about spaces.

\begin{exam}
	The following are some topological spaces.
	\begin{itm}
		\item $D^n = \{x \in \RR^n \mid d(x,0) \leq r\}$
		\item $S^{n-1} = \{x \in \RR^n \mid d(x,0) = r\}$
		\item $I = [0,1] \subset \RR$\
	\end{itm}
\end{exam}

1.2 continuous maps and homeomorphisms

\begin{defn}
	A map $f: X \to Y$ of sets is \textbf{continuous} if $f\inv = \{x \in X | f(x) \in V\}$ is open for all $V \subset Y$ open.
\end{defn}

Continuous maps are the natural maps between topological spaces.
If we talk about a map between topological spaces it is assumed to be continuous.

\begin{exam}
	Let $S \subset X$ and let $\iota: S \injto X$ be the inclusion.
	Let $V \subset X$ be open.
	Then $\iota\inv[V] = S \cap V$ which is open, hence $\iota$ is continuous.
\end{exam}

\begin{rmk}
	The composition of continuous maps is continuous.
\end{rmk}

\begin{rmk}
	Topological spaces form a category $\Top$.
	What are the morphisms in this category?
	When are two spaces the same topologically?
\end{rmk}

\begin{defn}
	A map $f: X \to Y$ is a \textbf{homeomorphism} if
	\begin{itm}
		\item $f$ is continuous
		\item $f$ is bijective
		\item $f\inv$ is continuous
	\end{itm}
	Alternatively, a homeomorphism is a bijection of sets where $f[U]$ is open in $Y$ if and only if $U$ is open in $X$.
	If a homeomorphism exists then we write $X \cong Y$.
\end{defn}

In topology we consider to be the same some spaces which would not be considered the same in analysis or geometry.

\begin{exam}
	The unit square is isomorphic to the unit disc.
	\[D^n \cong \square^n = \{x \in \RR^n | -1 \leq x_i \mid \leq 1\}\]
	The boundary of $\square^n$ is not smooth but this does not matter in topology.
\end{exam}

1.3 Elementary operations on topological spaces

Let $X$ and $Y$ be topological spaces.

\begin{defn}
	The \textbf{product space} $X \times Y$ is:
	\begin{itm}
		\item the set $X \times Y = \{(x,y) \mid x \in X, y \in Y\}$, and where
		\item open sets in $X \times Y$ are the unions of $U \times V$ where $U$ is open in $X$ and $V$ is open in $Y$
	\end{itm}
	(This is the product in the category $\Top$).
\end{defn}

We have (continuous) maps
\begin{itm}
	\item $P_X: X \times Y \to X, (x,y) \mapsto x$
	\item $P_X: X \times Y \to Y, (x,y) \mapsto y$
\end{itm}

\begin{exer}
	Let $f: Z \to X \times Y$.
	Then $f$ is continuous if and only if $P_Xf$ and $P_Yf$ is continuous.
\end{exer}

\begin{exam}
	Here are some examples of product spaces.
	\begin{itm}
		\item $\RR \times \RR \cong \RR^2$
		\item $S^1 \times S^1 \cong T^2$ (the torus)
	\end{itm}
\end{exam}

\begin{defn}
	Take an equivalence relation $\sim$ on $X$.
	Define $Z \defeq X/\sim$ be the set of equivalence classes with quotient map $q: X \to Y$.
	The \textbf{quotient topology} on $Y$ is the topology where the open sets are the sets of the form
	\[\{U \subset Y \mid q\inv[U] \in \tau\}\]
\end{defn}

\begin{exam}
	(Quotient by a subspace)
	Let $S \subset X$.
	Then for $x \neq y \in X$, let $x\sim y$ if and only if $x,y \in S$.
	Then $X/S \defeq X/\sim$.
\end{exam}

\begin{defn}
	Consider the (disjoint) \textbf{union} of sets $X \amalg Y$. (This is the coproduct in $\Top$.)
	The open subsets are $U \amalg V$ for $U$ open in $X$ and $V$ open in $Y$.
	We have inclusion maps $\iota_X: X \to X \amalg Y, x \mapsto x$ and $\iota_Y: Y \to X \amalg Y, y \mapsto y$.
\end{defn}

\begin{exer}
	Let $f: X \amalg Y \to Z$ be a map of sets.
	Then $f$ is continuous if and only if $f\iota_X$ and $f\iota_Y$ are continuous.
\end{exer}
