\section{2018-04-18 Lecture}

\begin{defn}
	For a space $X$ define a set
	\[\wt X^0 = \{[\gamma] \mid \gamma \text{ a path in } X \text{ starting at } x_0\}.\]
	Then there is a map of sets
	\begin{align*}
		p: \wt X^0 &\to X \\
		[\gamma] &\mapsto \gamma(1)
	\end{align*}
\end{defn}

\begin{prop}
	If $X$ is PLPC and semilocally simply connected, then for $H \leq \pi_1(X,x_0)$ there exists a connected covering space $p:(X_H,\wt x_0) \to (X,x_0)$ with $p_*(\pi_1(X,x_0))=H$.
\end{prop}

\begin{proof}
	We first consider the case $H=1$.
	Then we need $\wt X$ to be simply connected.
	Define a basis for the topology on $X$ by
	\[\ca U = \{U \subseteq X \mid U \text{ is path connected and } \iota_*: \pi_1(U) \to \pi_1(X) \text{ is trivial}\}\]
	We now check that this is a basis for the topology.
	
	We need to show that every open set in $X$ is a union of the basis elements.
	Let $x \in W \subseteq X$ where $W$ is any open subset of $X$.
	Then since $X$ is semilocally simply connected, there exists a neighbourhood $V$ of $x$ with the homomorphism induced by the inclusion being trivial.
	Then since $X$ is locally path connected there exists an open set $U^x$ with $x \in U^x \subseteq V \cap W$ with $U^x$ path connected.
	Then the homomorphism induced by the inclusion $U^x \injto X$ is trivial, so $U^x \in \ca U$.
	Then by construction, $W = \bigcup_{x \in W} U^x$.
	
	Now let $U \in \ca U$ and let $\gamma$ be a path in $X$ from $x_0$ to a point in $U$.
	Then define
	\[U_{[\gamma]} = \{[\gamma\cdot\eta] \mid \eta \text{ is a path in } U \text{ starting at } \gamma(1)\} \subseteq \wt X^0\]
	Now we have the following properties:
	\begin{itm}
		\io $U_{[\gamma]}$ depends only on $[\gamma]$.
		\io $p: Y_{[\gamma]} \to U$ is a bijection.
		\io If $[\gamma'] \in U_{[\gamma]}$, then $U_{[\gamma]} = U_{[\gamma']}$.
	\end{itm}
	Then $\{U_{[\gamma]}\}_\gamma$ forms a basis for a topology on $\wt X^0$.
	Clearly the basis elements cover $\wt X^0$.
	We will omit the proof of the intersection property.
	We will now write $\wt X$ for $\wt X^0$ to indicate that it is a topological space.
	
	Note that $p: U_{[\gamma]} \to U$ is a homeomorphism.
	Then $p: \wt X \to X$ is continuous and is a covering space.
	
	We now show that $\wt X$ is path connected.
	Take $[\gamma] \in \wt X$.
	Then for $y \in I$, define $\gamma_t: I \to X$ by
	\begin{align*}
		\gamma_t(s) =
		\begin{cases}
			\gamma(s) & s \leq t \\
			\gamma(t) & s>t
		\end{cases}
	\end{align*}
	Then $t \mapsto [\gamma_t]$ is a path from $[c]$ to $[\gamma]$ in $\wt X$.
	
	We now show that $\wt X$ is simply connected.
	Take a loop $\gamma in X$ at $x_0$.
	Then $\wt \gamma: t \mapsto [\gamma_t]$ is also a loop with $[\gamma]=[c]$.
	Thus $\gamma$ is null-homotopic.
	
	We now consider the general case.
	Define
	\[X_H = \wt X/\sim \quad \text{where} \quad [\gamma_1]\sim[\gamma_2] \iff [\gamma_1\cdot\ol\gamma_2] \in H\]
	Then $p$ factors through $H$ to give $p_H: X_H \to X$ which is a covering space.
\end{proof}