\section{2018-04-23 Lecture}

We know about the correspondence between connected covering spaces with a basepoint and subgroups of the fundamental group.
What if we ignore basepoints?

\begin{rmk}
  Consider the covering spaces
  \begin{align*}
    p: (\wt X,\wt x_0) &\to (X,x_0) \\
    p: (\wt X,\wt y_0) &\to (X,x_0) 
  \end{align*}
  and take a path $\wt \gamma$ in $\wt X$ from $\wt x_0$ to $\wt y_0$.
  Now $\gamma = p\wt\gamma$ is a loop in $X$ at $x_0$.
  Then the following diagram commutes:
  \begin{equation*}
    \begin{tikzcd}
      \pi_1(\wt X,\wt x_0) \ar[r, "p_*"] & \pi_1(X,x_0) \\
      \pi_1(\wt X,\wt y_0) \ar[u, "\beta_{\wt\gamma}"] \ar[r, "p_*"] & \pi_1(X,x_0) \ar[u, "\beta_\gamma"]
    \end{tikzcd}
  \end{equation*}
  where the vertical arrows are isomorphisms.
  Then $\beta_\gamma(g) = hgh\inv$ for any $g \in \pi_1(X,x_0)$.
\end{rmk}

\begin{thm}
  Let $X$ be PLPC and semi-locally simply connected.
  Then there is a one-to-one correspondence between isoclasses of connected $p: \wt X \to X$ and conjugacy classes of subgroups of $\pi_1(X,x_0)$.
\end{thm}

3.4 Deck transformations of $p: \wt X \to X$

\begin{defn}
  Let $G(\wt X)$ be the group of automorphisms of the covering space $p:\wt X \to X$.
  These are the homeomorphisms which are isomorphisms from $p: \wt X \to X$ to itself, that is, the homeomorphisms $\phi: \wt X \to \wt X$ satisfiying $p=p\phi$.
  These are the \textbf{deck transformations} of the covering space $p: \wt X \to X$.
\end{defn}

\begin{defn}
  A covering space $p: \wt X \to X$ is called \textbf{normal} if for each $x \in X$ and $\wt x_1, \wt x_2 \in p\inv[x]$ there is a $\phi \in G(\wt X)$ with $\phi(\wt x_1)=\wt x_2$.
\end{defn}

\begin{exer}
  Show that we can equivalently replace ``for each $x \in X$'' with ``for one $x_0 \in X$'' in the above definition.
\end{exer}

\begin{prop}
  Let $p:(\wt X,\wt x_0) \to (X,x_0)$ be a covering space with $\wt X$ connected.
  Define $H = p_*(\pi_1(\wt X,\wt x_0))$ and $G = \pi_1(X,x_0)$.
  Then
  \begin{enum}
    \io $p$ is normal $\iff H \trianglelefteq G$.
    \io $G(\wt X) = N_G(H)/H$ where $N_G(H)$ is the normaliser of $G$ in $H$.
    \io $p$ is normal $\iff G(\wt X) = G/H$.
    \io $\wt X$ is simply connected $ \iff G(\wt X)=G$.
  \end{enum}
\end{prop}

\begin{proof}
  See lecture notes.
\end{proof}

\begin{exam}
  Consider the covering space of $S^1 \vee S^1$ given by the infinite unit grid in $\RR^2$.
\end{exam}

3.5 Group actions

\begin{defn}
  An action of a group $G$ on a topological space $Y$ is a group homomorphism $\rho: G \to \Homeo(Y)$.
  We omit the $\rho$ in notation.
  
  Define $Y/G = Y/\sim$ where we define $y_1 \sim y_2$ if $gy_1=y_2$ for some $g \in G$.
\end{defn}

\begin{defn}
  $\rho: G \to \Homeo(Y)$ is a \textbf{covering space action} if each $y \in Y$ has an open neighbourhood $V$ such that $g_1[V] \cap g_2[V] \neq \emptyset$ implies that $g_1=g_2$.
\end{defn}

\begin{prop}
  \leavevmode
  \begin{enum}
    \io For a covering space $p: \wt X \to X$ with $\wt X$ connected, the inclusion $G(\wt X) \injto \Homeo(\wt X)$ is a covering space action.
    \io For a covering space action $G \to \Homeo(Y)$:
      \begin{enum}
	\io $q: Y \to Y/G$ is a normal covering space.
	\io $G(Y) \cong G$ is $Y$ is connected.
      \end{enum}
  \end{enum}
\end{prop}
